---
title: "Adv Candidate - SVM"
author: "Group 10"
date: "2020/4/1"
output: html_document
---
### Result:   
|  svm model |  
| ------ | ------ | 
| train accuracy | 68.55% | 
| test accuracy | 43.2% | 
---

```{r message=FALSE}
if(!require("EBImage")){
  source("https://bioconductor.org/biocLite.R")
  biocLite("EBImage")
}
if(!require("R.matlab")){
  install.packages("R.matlab")
}
if(!require("readxl")){
  install.packages("readxl")
}
if(!require("dplyr")){
  install.packages("dplyr")
}
if(!require("readxl")){
  install.packages("readxl")
}
if(!require("ggplot2")){
  install.packages("ggplot2")
}
if(!require("caret")){
  install.packages("caret")
}
if(!require("gbm")){
  install.packages("gbm")
}
if(!require("h2o")){
  install.packages("h2o")
}
if(!require("e1071")){
  install.packages("e1071")
}
if(!require("kernlab")){
  install.packages("kernlab")
}
library(R.matlab)
library(readxl)
library(dplyr)
library(EBImage)
library(ggplot2)
library(caret)
library(gbm)
library(h2o)
library(e1071)
library(kernlab)
```

### Step 0 set work directories
```{r wkdir, eval=FALSE}
set.seed(0)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
# here replace it with your own path or manually set it in RStudio to where this rmd file is located. 
# use relative path for reproducibility
```

Provide directories for training images. Training images and Training fiducial points will be in different subfolders. 
```{r}
train_dir <- "../../data/train_set/" # This will be modified for different data sets.
train_image_dir <- paste(train_dir, "images/", sep="")
train_pt_dir <- paste(train_dir,  "points/", sep="")
train_label_path <- paste(train_dir, "label.csv", sep="") 
```

### Step 1: set up controls for evaluation experiments.


```{r exp_setup}
run.cv=TRUE # run cross-validation on the training set
K <- 5  # number of CV folds
run.feature.train=TRUE # process features for training set
run.test=TRUE # run evaluation on an independent test set
run.feature.test=TRUE # process features for test set
```

### Step 2: import data and train-test split 
```{r}
#train-test split
info <- read.csv(train_label_path)
n <- nrow(info)
n_train <- round(n*(4/5), 0)
train_idx <- sample(info$Index, n_train, replace = F)
test_idx <- setdiff(info$Index,train_idx)
```


Fiducial points are stored in matlab format. In this case, we select inner face points as fiducial points, 33 points in total. In this step, we read them and store them in a list.
```{r read fiducial points}
#function to read fiducial points
#input: index
#output: matrix of fiducial points corresponding to the index
readMat.matrix <- function(index){
     return(round(readMat(paste0(train_pt_dir, sprintf("%04d", index), ".mat"))[[1]],0))
}
#load fiducial points
n_files <- length(list.files(train_image_dir))
fiducial_pt_list <- lapply(1:n_files, readMat.matrix)
save(fiducial_pt_list, file="../output/fiducial_pt_list.RData")
```

```{r select fiducial_pt}
#inner points
for (i in 1:2500){
  fiducial_pt_list[[i]] = fiducial_pt_list[[i]][c(19,21,23, 27,29,31,2,4,6,8,11,13,15,17,1,10,41,42,44,46,47,50,51,38,53,54,55,57,59,62,67,71,75),]
}
```
### Step 3: construct features and responses


```{r feature}
source("../../lib/feature.R")
tm_feature_train <- NA
if(run.feature.train){
  tm_feature_train <- system.time(dat_train <- feature(fiducial_pt_list, train_idx))
}

tm_feature_test <- NA
if(run.feature.test){
  tm_feature_test <- system.time(dat_test <- feature(fiducial_pt_list, test_idx))
}
save(dat_train, file="../output/feature_train.RData")
save(dat_test, file="../output/feature_test.RData")

```

### Step 4: Advanced model candidate - SVM 


```{r svm}
# tuning, pick the best parameters 
classifier = train(form = emotion_idx~., data =dat_train, method = 'svmRadial')
save(classifier,file="../output/svmclassifier.RData")

classifier$bestTune
#training
model_svm = ksvm(emotion_idx ~ ., data = dat_train, kernel = 'rbfdot', C = 1, sigma = 0.0006176977)

ksvm.pred.train <- predict(object = model_svm)
accu <- mean(dat_train$emotion_idx == ksvm.pred.train)

# test
ksvm.pred.test <- predict(object = model_svm, newdata = dat_test)
accu.test <- mean(dat_test$emotion_idx == ksvm.pred.test)

```

###Reference
- https://www.kdnuggets.com/2016/06/select-support-vector-machine-kernels.html
- https://www.hackerearth.com/blog/developers/simple-tutorial-svm-parameter-tuning-python-r/
