---
title: "Adv Candidate - PCA+LDA"
author: "Group 10"
output:
  html_document:
    df_print: paged
---
### Result:   
|  pca lda model |  
| ------ | ------ | 
| train accuracy | 56.85% | 
| test accuracy | 47.2% | 
| training time | 16.621s |
| testing time | 0.005s |


```{r message=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)){
    install.packages("BiocManager")
    BiocManager::install()
    BiocManager::install("EBImage")
}
if(!require("R.matlab")){
  install.packages("R.matlab")
}
if(!require("readxl")){
  install.packages("readxl")
}

if(!require("dplyr")){
  install.packages("dplyr")
}
if(!require("readxl")){
  install.packages("readxl")
}

if(!require("ggplot2")){
  install.packages("ggplot2")
}

if(!require("caret")){
  install.packages("caret")
}

if(!require("gbm")){
  install.packages("gbm")
}
if(!require("MASS")){
  install.packages("MASS")
}

library(R.matlab)
library(readxl)
library(dplyr)
library(EBImage)
library(ggplot2)
library(caret)
library(gbm)
library(MASS)
```

### Step 0 set work directories
```{r directories, eval=FALSE}
# # set to the local path
# #setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
# setwd("/Users/sqzhao/Documents/columbia/stat5243\ ads/project3/Spring2020-Project3-group10/doc/candidate\ models/Main_pca.Rmd")

train_dir <- "../../data/train_set/"
train_image_dir <- paste(train_dir, "images/", sep="") #"../../data/train_set/images/"
train_pt_dir <- paste(train_dir,  "points/", sep="") #"../../data/train_set/points/"
train_label_path <- paste(train_dir, "label.csv", sep="") #"../../data/train_set/label.csv"
```

### Step 1: set up controls for evaluation experiments.
```{r setup contol harness}
run.cv <- TRUE # run cross-validation on the training set
K <- 5  # number of CV folds
run.feature.train <- TRUE # process features for training set
run.feature.test <- TRUE # process features for test set
run.test <- TRUE # run evaluation on an independent test set
```

### Step 2: import data and train-test split 
```{r train-test split data}
#train-test split
#info <- read.csv(train_label_path)
info <- read.csv("/Users/sqzhao/Documents/columbia/stat5243\ ads/project3/Spring2020-Project3-group10/data/train_set/label.csv") 
n <- nrow(info) 
n_train <- round(n*(4/5), 0) #2000 images
set.seed(1)
train_idx <- sample(info$Index, n_train, replace = F)
test_idx <- setdiff(info$Index,train_idx) #500 images
```


### Step 3: construct features and responses

```{r}
load("../../output/fiducial_pt_list.RData")
fiducial_pt_half <- list()
# all points:
# middle: 35:44 nose
#         50:52,56:59,62,63 mouse
# left: 1:9,19:26 eye and brow
# __________
# contour of face: 64:71
for (i in 1:2500){
  fiducial_pt_half[[i]] = fiducial_pt_list[[i]][c(1:9,19:26,35:44,50:52,56:59,62,63,64:71),]
}

fiducial_pt_inner <- list()
# inner: 
for (i in 1:2500){
  #fiducial_pt_inner[[i]] <- fiducial_pt_list[[i]][c(19,21,23, 27,29,31,2,4,6,8,11,13,15,17,1,10,41,42,44,46,47,50,51,38,53,54,55,57,59,62,67,71,75),]
  fiducial_pt_inner[[i]] <- fiducial_pt_list[[i]][c(2:9, 11:23, 27:31, 37, 39:49,50,52,54,56),]
}

```


```{r inner train-test split}
source("../../lib/feature.R")
tm_feature_train <- NA # need add time of construction later
tm_feature_test <- NA

if(run.feature.train){

 dat_train_half <- feature(fiducial_pt_inner, train_idx)
}

if(run.feature.train){
  dat_test_half <- feature(fiducial_pt_inner, test_idx)}


```

Step 4: Advanced Model
PCA for dimension reduction of the features and resolve the colllinearity problem when fitting to LDA.

Train and Test
```{r pca}
train_start = proc.time()
pca <- princomp(dat_train_inner[,-ncol(dat_train_inner)], cor = TRUE)

# Calculate variability of each component
pr.cvar <- pca$sdev ^ 2
# Variance explained by each principal component: pve
pve_cov <- pr.cvar/sum(pr.cvar)
plot(pve_cov[1:100], xlab = "Principal Component", ylab = "Proportion of Variance Explained", ylim = c(0, 1), type = "b")
# Plot cumulative proportion of variance explained
plot(cumsum(pve_cov[1:100]), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 1), type = "b")
abline(h=0.95, col="red", lty=3)
abline(v=40, col="red", lty=3)
abline(v=50, col="red", lty=3)


n.pc <- 45
train_x_pca <- data.frame(pca$scores[,1:n.pc]) 
train_y <- dat_train_inner[ncol(dat_train_inner)]
train_pca <- cbind(train_x_pca, train_y)

train <- train_pca

lda.model <- lda(emotion_idx ~ ., data=train)
lda.train.pred <-  predict(lda.model, train[-(n.pc+1)])
confusionMatrix(lda.train.pred$class, train$emotion_idx)$overall["Accuracy"]

train_end = proc.time()
train_time <- train_end-train_start
train_time

test_x_pca <- data.frame(predict(pca, dat_test_inner[,-1893])[,1:n.pc])
test_y <- dat_test_inner[ncol(dat_test_inner)]
test_pca <- cbind(test_x_pca, test_y)
test <- test_pca

test_time <- system.time(lda.test.pred <-  predict(lda.model, test[-(n.pc+1)]))
test_time
confusionMatrix(lda.test.pred$class, test$emotion_idx)$overall["Accuracy"]
```


### reference:
- methodology of advanced model: https://link.springer.com/chapter/10.1007%2F978-3-030-16848-3_11
- explained feature extraction(PCA, LDA): https://medium.com/@cxu24/common-methods-for-feature-extraction-pca-and-lda-7b1f5679e3bf
http://ce.sharif.edu/courses/91-92/2/ce725-2/resources/root/Lectures/Feature%20Extraction.pdf
- LDA as a classifier or served to dimension reduction https://towardsdatascience.com/is-lda-a-dimensionality-reduction-technique-or-a-classifier-algorithm-eeed4de9953a, https://www.kaggle.com/shravank/predicting-breast-cancer-using-pca-lda-in-r
